---
title: "08_02_2020"
author: "Evelyn Cai"
date: "8/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Load libraries
library(tidyverse)
library(janitor)
library(skimr)
library(lubridate)
library(tidymodels)
library(readxl)
library(patchwork)
library(ggridges)

# Read in data and clean names/columns
df <- read_csv("urbana_traffic_stops.csv",
               col_types= cols(
                 `AGE AT CONTACT` = col_character())) %>%
  clean_names() %>%
  mutate(date_of_contact = mdy(date_of_contact),
         age_at_contact = case_when(age_at_contact == "NN" ~ NA_character_,
                                    TRUE ~ age_at_contact),
         age_at_contact = as.numeric(age_at_contact))

# Read in duration CSV
durations <- read_xlsx("Stop Duration thru 2017.xlsx") %>%
  clean_names()

# Read in accident data
accidents <- read_xlsx("Drivers in Accidents - Benchmarks.xlsx")
```

```{r ages_deprecated, eval=FALSE, echo=FALSE}
# Below, I explored the differences in the two datasets, since joining seems to yield some extra data points. It seems there are 130 incident numbers in `ages` that is not in `df` and 1558 incident numbers in `df` that is not in `ages`. Therefore, I will choose to ignore the age piece right now.
# only_ins <- df %>% select(incident_number) %>% unique()
# only_age_ins <- ages %>% select(incident_number) %>% unique()
# anti_join(only_age_ins, only_ins)

# Read in additional ages
ages <- read_xlsx("Additional IDOT info.xlsx") %>%
  clean_names() %>%
  rename("incident_number" = "incident") %>%
  select(incident_number, age_at_contact) %>%
  unique()

# Join for ages
tstops <- left_join(df, ages, by = "incident_number") %>%
  rename("age_at_contact" = "age_at_contact.y") %>%
  mutate(age_at_contact = case_when(is.na(age_at_contact.x) ~ age_at_contact,
                                    TRUE ~ age_at_contact.x)) %>%
  select(-age_at_contact.x)


# Below, I explored the differences in the two datasets, since joining seems to yield some extra data points. It seems there are 130 incident numbers in `ages` that is not in `df` and 1558 incident numbers in `df` that is not in `ages`. Therefore, I will choose to ignore the age piece right now.

 only_ins <- df %>% select(incident_number) %>% unique()
 only_age_ins <- ages %>% select(incident_number) %>% unique()
 anti_join(only_age_ins, only_ins)

```

```{r cleaning_df, echo=FALSE}
tstops <- df %>%
  group_by(incident_number) %>%
  # Create new variables that collapse multiple offenses into one column
  mutate(crime_descs  = paste(crime_description, collapse =", "),
         crime_codes = paste(crime_code, collapse = ", "),
         crime_status_descs = paste(crime_status_description, collapse = ", "),
         crime_categories = paste(crime_category, collapse = ", "),
         crime_category_descs = paste(crime_category_description, collapse = ", "),
         # Create new variable to count the number of offenses
         number_of_crimes = length(crime_codes)) %>%
  # ungroup
  ungroup() %>%
  # Deselect duplicate columns
  select(-crime_code, -crime_description, -crime_status_description, -crime_category, -crime_category_description) %>%
  # Filter for unique incidents
  unique()

## Create a new dataset that contains demographic information

tstops_demo <- tstops %>%
  # Select only for useful demographic and logistical information
  select(date_of_contact, time_of_contact, incident_number,
        motivation_for_stop_desc, race_description, sex_description,
        age_at_contact, residency_code, reason_for_stop,
        type_of_violation, vehicle_year, crime_descs, crime_status_descs, crime_category_descs) %>%
  # Extract year and month from date
  mutate(year = year(date_of_contact),
         month = month(date_of_contact),
         # Change "Caucasian" to "White" and factorize race
         race_description = as.factor(case_when(
           race_description == "CAUCASIAN" ~ "WHITE",
           race_description == "AFRICAN AMERICAN" ~ "BLACK",
           TRUE ~ race_description)))
```

```{r echo=FALSE, fig.height=6, fig.width=10}
# Graphing race by year

race_by_year_unique <- tstops_demo %>%
  group_by(race_description, year) %>%
  mutate(race_ct = n()) %>%
  ungroup() %>%
  select(race_description, race_ct, year) %>%
  unique()

## Todo: Clean up x labels on race_by_year1 and have years increase by 1; remove x labels for second graph

race_by_year1 <- race_by_year_unique %>%
  ggplot(aes(x = year, y = race_ct ,fill = race_description)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(fill = "Race",
       title = "Number of stops by race per year",
       caption = "Data Publicly Available from Urbana Police Department") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  scale_fill_brewer(palette = "PRGn") +
  xlab("Year") +
  ylab("Number of Stops") +
  scale_x_continuous(breaks = c(2012, 2013, 2014, 2015, 2016,2017, 2018, 2019, 2020))
  
race_by_year2 <- race_by_year_unique %>%
  ggplot(aes(x = year, y = race_ct ,fill = race_description)) +
  geom_col() +
  labs(fill = "Race",
       title = "Number of stops by race per year",
       caption = "Data Publicly Available from Urbana Police Department") +
  theme_minimal() +
  theme(legend.position = "bottom",
        axis.text.x = element_blank()) +
  scale_fill_brewer(palette = "PRGn") +
  xlab("Year") +
  ylab("Number of Stops") +
  facet_grid(~race_description)

# Graph

race_by_year1 + race_by_year2

```

## Disparity Ratio Calculations

```{r disparity-ratios, fig.height=3, fig.width=8}
disparity_ratios <- tstops_demo %>%
  filter(!race_description %in% c("AMERICAN INDIAN/ALASKAN", "UNKNOWN")) %>%
  group_by(year) %>%
  mutate(all = n()) %>%
  ungroup() %>%
  group_by(year, race_description) %>%
  mutate(race_ct = n(),
         race_percentage = 100 * race_ct / all,
         disparity_ratio = case_when(
           year >= 2018 & race_description == "ASIAN" ~ race_percentage / 8.37,
           year >= 2018 & race_description == "BLACK" ~ race_percentage / 22.26,
           year >= 2018 & race_description == "HISPANIC" ~ race_percentage / 4.84,
           year >= 2018 & race_description == "WHITE" ~ race_percentage / 62.23,
           year == 2017 & race_description == "ASIAN" ~ race_percentage / 9.07,
           year == 2017 & race_description == "BLACK" ~ race_percentage / 20.36,
           year == 2017 & race_description == "HISPANIC" ~ race_percentage / 4.4,
           year == 2017 & race_description == "WHITE" ~ race_percentage / 63.7,
           year == 2016 & race_description == "ASIAN" ~ race_percentage / 9.91,
           year == 2016 & race_description == "BLACK" ~ race_percentage / 18.56,
           year == 2016 & race_description == "HISPANIC" ~ race_percentage / 4.04,
           year == 2016 & race_description == "WHITE" ~ race_percentage / 65.54,
           year <= 2015 & race_description == "ASIAN" ~ race_percentage / 9.96,
           year <= 2015 & race_description == "BLACK" ~ race_percentage / 18.07,
           year <= 2015 & race_description == "HISPANIC" ~ race_percentage / 3.52,
           year <= 2015 & race_description == "WHITE" ~ race_percentage / 67.26)) %>%
  ungroup() %>%
  group_by(year, race_description, sex_description) %>%
  mutate(race_gender_ct = n(),
         race_gender_percentage = 100 * race_gender_ct / all,
         disparity_ratio_gender = case_when(
           year >= 2018 & race_description == "ASIAN" ~ race_gender_percentage / 8.37,
           year >= 2018 & race_description == "BLACK" ~ race_gender_percentage / 22.26,
           year >= 2018 & race_description == "HISPANIC" ~ race_gender_percentage / 4.84,
           year >= 2018 & race_description == "WHITE" ~ race_gender_percentage / 62.23,
           year == 2017 & race_description == "ASIAN" ~ race_gender_percentage / 9.07,
           year == 2017 & race_description == "BLACK" ~ race_gender_percentage / 20.36,
           year == 2017 & race_description == "HISPANIC" ~ race_gender_percentage / 4.4,
           year == 2017 & race_description == "WHITE" ~ race_gender_percentage / 63.7,
           year == 2016 & race_description == "ASIAN" ~ race_gender_percentage / 9.91,
           year == 2016 & race_description == "BLACK" ~ race_gender_percentage / 18.56,
           year == 2016 & race_description == "HISPANIC" ~ race_gender_percentage / 4.04,
           year == 2016 & race_description == "WHITE" ~ race_gender_percentage / 65.54,
           year <= 2015 & race_description == "ASIAN" ~ race_gender_percentage / 9.96,
           year <= 2015 & race_description == "BLACK" ~ race_gender_percentage / 18.07,
           year <= 2015 & race_description == "HISPANIC" ~ race_gender_percentage / 3.52,
           year <= 2015 & race_description == "WHITE" ~ race_gender_percentage / 67.26)) %>%
  ungroup() %>%
  select(year, race_description, disparity_ratio, disparity_ratio_gender, sex_description, race_percentage) %>%
  unique()

## Only 2012 is off compared to the 2018 report, so I omitted it

# Graph disparity ratios
disparity_ratios %>%
  ggplot(aes(year, disparity_ratio, color = race_description)) +
  geom_line() +
  scale_x_continuous(breaks = c(2013, 2014, 2015, 2016,2017, 2018, 2019, 2020)) +
  scale_y_continuous(breaks = seq(0, 2.5, by = 0.5)) +
  theme_minimal() +
  labs(title = "Disparity Ratios",
       caption = "Data Publicly Available from Urbana Police Department") +
  scale_color_brewer(palette = "Dark2") +
  xlab("Year") +
  ylab("Disparity Ratio") +
  geom_hline(yintercept = 1, color = "red")
```

```{r fig.height=4.5, fig.width=8}
# Graph disparity ratios by gender
disparity_ratios %>%
  ggplot(aes(year, disparity_ratio_gender, color = sex_description)) +
  geom_line() +
  scale_x_continuous(breaks = c(2013, 2014, 2015, 2016,2017, 2018, 2019, 2020)) +
  scale_y_continuous(breaks = seq(0, 2.5, by = 0.5)) +
  theme_minimal() +
  labs(title = "Disparity Ratios by Gender",
       caption = "Data Publicly Available from Urbana Police Department") +
  geom_hline(yintercept = 0.5, color = "red") +
  facet_wrap(~race_description)

## To do gender*race disparity ratios: We'd need to either 1. Account for the different frequencies in male vs females being stopped, or 2. Find out the average # males and females stopped in Urbana to use as a "checkpoint" or baseline. The above chart displays disparity ratios by gender, which assumes that there are an equal number of men and women driving on the streets.
```

```{r time_of_contact, fig.width = 10, fig.height=6}
tstops_demo %>%
  filter(!race_description %in% c("AMERICAN INDIAN/ALASKAN", "UNKNOWN")) %>%
  mutate(time_hours = as.numeric(time_of_contact, units = "secs"),
         time_hours = time_hours/3600,
         time_of_day = factor(case_when(time_hours >= 0 & time_hours <= 6 ~ "Midnight to 6am",
                                 time_hours > 6 & time_hours <= 12 ~ "6am to noon",
                                 time_hours > 12 & time_hours <= 18 ~ "Noon to 6pm",
                                 time_hours > 18 & time_hours <= 24 ~ "6pm to Midnight",
                                 TRUE ~ NA_character_),
                              levels = c("Midnight to 6am", "6am to noon", "Noon to 6pm", "6pm to Midnight"))) %>%
  group_by(year) %>%
  mutate(all = n()) %>%
  ungroup() %>%
  group_by(year, race_description, time_of_day) %>%
  mutate(race_time_ct = n(),
         race_time_percentage = 100 * race_time_ct / all,
         disparity_ratio_time = case_when(
           year >= 2018 & race_description == "ASIAN" ~ race_time_percentage / 8.37,
           year >= 2018 & race_description == "BLACK" ~ race_time_percentage / 22.26,
           year >= 2018 & race_description == "HISPANIC" ~ race_time_percentage / 4.84,
           year >= 2018 & race_description == "WHITE" ~ race_time_percentage / 62.23,
           year == 2017 & race_description == "ASIAN" ~ race_time_percentage / 9.07,
           year == 2017 & race_description == "BLACK" ~ race_time_percentage / 20.36,
           year == 2017 & race_description == "HISPANIC" ~ race_time_percentage / 4.4,
           year == 2017 & race_description == "WHITE" ~ race_time_percentage / 63.7,
           year == 2016 & race_description == "ASIAN" ~ race_time_percentage / 9.91,
           year == 2016 & race_description == "BLACK" ~ race_time_percentage / 18.56,
           year == 2016 & race_description == "HISPANIC" ~ race_time_percentage / 4.04,
           year == 2016 & race_description == "WHITE" ~ race_time_percentage / 65.54,
           year <= 2015 & race_description == "ASIAN" ~ race_time_percentage / 9.96,
           year <= 2015 & race_description == "BLACK" ~ race_time_percentage / 18.07,
           year <= 2015 & race_description == "HISPANIC" ~ race_time_percentage / 3.52,
           year <= 2015 & race_description == "WHITE" ~ race_time_percentage / 67.26)) %>%
  ungroup() %>%
  select(year, race_description, disparity_ratio_time, sex_description, race_time_percentage, time_of_day) %>%
  unique() %>%
  ggplot(aes(year, disparity_ratio_time, color = time_of_day)) +
  geom_line() +
  scale_x_continuous(breaks = c(2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020)) +
  scale_y_continuous(breaks = seq(0, 2.5, by = 0.5)) +
  theme_minimal() +
  labs(title = "Disparity Ratios by Time of Day",
       caption = "Data Publicly Available from Urbana Police Department") +
  facet_wrap(~race_description)

## Do it again to follow sunrise/sunset

tstops_demo %>%
  filter(!race_description %in% c("AMERICAN INDIAN/ALASKAN", "UNKNOWN")) %>%
  mutate(time_hours = as.numeric(time_of_contact, units = "secs"),
         time_hours = time_hours/3600,
         time_of_day = factor(case_when(time_hours >= 0 & time_hours <= 7 ~ "Night",
                                        time_hours >= 7 & time_hours <= 20 ~ "Day",
                                        time_hours >= 20 ~ "Night",
                                        TRUE ~ NA_character_),
                              levels = c("Day", "Night"))) %>%
  group_by(year) %>%
  mutate(all = n()) %>%
  ungroup() %>%
  group_by(year, race_description, time_of_day) %>%
  mutate(race_time_ct = n(),
         race_time_percentage = 100 * race_time_ct / all,
         disparity_ratio_time = case_when(
           year >= 2018 & race_description == "ASIAN" ~ race_time_percentage / 8.37,
           year >= 2018 & race_description == "BLACK" ~ race_time_percentage / 22.26,
           year >= 2018 & race_description == "HISPANIC" ~ race_time_percentage / 4.84,
           year >= 2018 & race_description == "WHITE" ~ race_time_percentage / 62.23,
           year == 2017 & race_description == "ASIAN" ~ race_time_percentage / 9.07,
           year == 2017 & race_description == "BLACK" ~ race_time_percentage / 20.36,
           year == 2017 & race_description == "HISPANIC" ~ race_time_percentage / 4.4,
           year == 2017 & race_description == "WHITE" ~ race_time_percentage / 63.7,
           year == 2016 & race_description == "ASIAN" ~ race_time_percentage / 9.91,
           year == 2016 & race_description == "BLACK" ~ race_time_percentage / 18.56,
           year == 2016 & race_description == "HISPANIC" ~ race_time_percentage / 4.04,
           year == 2016 & race_description == "WHITE" ~ race_time_percentage / 65.54,
           year <= 2015 & race_description == "ASIAN" ~ race_time_percentage / 9.96,
           year <= 2015 & race_description == "BLACK" ~ race_time_percentage / 18.07,
           year <= 2015 & race_description == "HISPANIC" ~ race_time_percentage / 3.52,
           year <= 2015 & race_description == "WHITE" ~ race_time_percentage / 67.26)) %>%
  ungroup() %>%
  select(year, race_description, disparity_ratio_time, sex_description, race_time_percentage, time_of_day) %>%
  unique() %>%
  ggplot(aes(year, disparity_ratio_time, color = time_of_day)) +
  geom_line() +
  scale_x_continuous(breaks = c(2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020)) +
  scale_y_continuous(breaks = seq(0, 2.5, by = 0.5)) +
  theme_minimal() +
  labs(title = "Disparity Ratios by Time of Day (Day and Night)",
       caption = "Data Publicly Available from Urbana Police Department") +
  facet_wrap(~race_description)
```


```{r age, fig.height=8, fig.width=10, message=FALSE, warning=FALSE}

# Age distribution of drivers in 2018, 2019, and 2020

tstops_demo %>%
  filter(!race_description %in% c("AMERICAN INDIAN/ALASKAN", "UNKNOWN"),
         !is.na(age_at_contact)) %>%
  ggplot(aes(x = age_at_contact, y = race_description)) +
  geom_density_ridges(show.legend = FALSE) +
  facet_wrap(~year) +
  labs(title = "Age Distribution of Drivers by Race",
       caption = "Data Publicly Available from the Urbana Police Department",
       xlab = "Age at Contact") +
  ylab(NULL)
```

```{r duration, echo=FALSE, fig.height=8, fig.width=10}

# Join duration data with tstop demographic information
duration_join <- left_join(tstops_demo, durations, by = "incident_number")

duration_join %>%
  filter(!race_description %in% c("AMERICAN INDIAN/ALASKAN", "UNKNOWN"),
         !is.na(duration_of_stop)) %>%
  group_by(year, race_description) %>%
  mutate(avg_duration = mean(duration_of_stop)) %>%
  ungroup() %>%
  ggplot(aes(x = year, y = avg_duration, color = race_description)) +
  geom_line()
```

As you can see from the graph above, average durations are higher for all POC groups than white folks. Hispanic folks are stopped the most on average, but let's view a boxplot to account for extremes.

```{r echo=FALSE, fig.height=8, fig.width=10}
duration_join %>%
  filter(!race_description %in% c("AMERICAN INDIAN/ALASKAN", "UNKNOWN"),
         !is.na(duration_of_stop)) %>%
  ggplot(aes(x = race_description, y = duration_of_stop)) +
  geom_boxplot() +
  facet_wrap(~year)

```

```{r echo=FALSE, fig.height=8, fig.width=10}
accidents <- accidents %>%
  mutate(year = year(Date)) %>%
  filter(!is.na(Age))

# Plot of age distribution
vertical <- accidents %>%
  filter(Race %in% c("W", "H", "A", "B")) %>%
  summarize(vertical = mean(Age))

vertical <- as.numeric(vertical)

accidents %>%
  filter(Race %in% c("W", "H", "A", "B")) %>%
  ggplot(aes(x = Age, y = Race)) +
  geom_density_ridges(show.legend = FALSE) +
  scale_y_discrete(labels = rev(c("White", "Hispanic", "Black", "Asian"))) +
  scale_x_continuous(breaks = seq(0, 100, by = 10)) +
  geom_vline(xintercept = vertical) +
  labs(title = "Age Distribution of Drivers Involved in Accidents",
       subtitle = "Mean age is about 39 years old")

# Plotting age distribution by at fault
accidents %>%
  filter(Race %in% c("W", "H", "A", "B")) %>%
  ggplot(aes(x = Age, y = Race)) +
  geom_density_ridges(show.legend = FALSE) +
  facet_wrap(~`At Fault`) +
  scale_y_discrete(labels = rev(c("White", "Hispanic", "Black", "Asian"))) +
  scale_x_continuous(breaks = seq(0, 100, by = 10)) +
  labs(title = "Age Distribution of Drivers Involved in Accidents",
       subtitle = "Faceted by whether the driver was at fault or not")

accidents %>%
  filter(Race %in% c("W", "H", "A", "B"),
         year >= 2014) %>%
  group_by(year, `At Fault`, Race) %>%
  mutate(mean_age = mean(Age)) %>%
  select(year, `At Fault`, Race, mean_age) %>%
  unique() %>%
  ggplot(aes(year, mean_age, color = Race)) +
  geom_line() +
  scale_x_continuous(breaks = c(2014, 2015, 2016, 2017, 2018, 2019)) +
  scale_y_continuous(breaks = seq(0, 80, by = 40)) +
  theme_minimal() +
  labs(title = "Mean Age of Drivers Involved in Accidents by Race",
       caption = "Faceted by whether the driver was at fault or not") +
  facet_wrap(~`At Fault`)

# Some proportion tables
## How many people are at fault of each race / gender group per year
accidents %>%
  filter(Race %in% c("W", "H", "A", "B")) %>%
  group_by(year, Race, Sex) %>%
  mutate(all = n()) %>%
  ungroup() %>%
  group_by(year, Race, Sex, `At Fault`) %>%
  mutate(at_fault = n(),
         prop_at_fault = at_fault / all) %>%
  ungroup() %>%
  select(Race,Sex, `At Fault`, year, prop_at_fault) %>%
  unique()
```

# Things I want to do:

* Interaction plots and parallel slopes models for variables such as:
    + stops & resultant arrest %s
    + duration & demographic variables
    + % more serious than traffic violations out of all stopped
    + Impact of location (?) and time of day on the stop
    + Interaction on time of day
    
# Questions

* Which variable reveals the outcome of the stop, and which outcomes are considered "more serious than traffic violations"? (Is it crime category description: traffic offenses?)
* Which variable, and which levels of that variable, reveal an arrest outcome for traffic stops?
* Where can I get the duration of each traffic stop?
* What are meaningful time segmentations? (morning/night, light/dark, rush hour/ non, etc)
* Anything you want to be looked at?
