---
title: "Civitas: Traffic Stops and Racial Disparities in Urbana, IL"
author: "Evelyn Cai"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# Load libraries
library(tidyverse)
library(janitor)
library(skimr)
library(lubridate)
library(tidymodels)
library(readxl)
library(patchwork)
library(ggridges)
library(rstanarm)

# Read in data and clean names/columns
df <- read_csv("urbana_traffic_stops.csv",
               col_types= cols(
                 `AGE AT CONTACT` = col_character())) %>%
  clean_names() %>%
  mutate(date_of_contact = mdy(date_of_contact),
         age_at_contact = case_when(age_at_contact == "NN" ~ NA_character_,
                                    TRUE ~ age_at_contact),
         age_at_contact = as.numeric(age_at_contact))

# Read in duration CSV
durations <- read_xlsx("Stop Duration thru 2017.xlsx") %>%
  clean_names()

# Read in accident data
accidents <- read_xlsx("Drivers in Accidents - Benchmarks.xlsx")
```

```{r ages_deprecated, eval=FALSE, echo=FALSE}
# Below, I explored the differences in the two datasets, since joining seems to yield some extra data points. It seems there are 130 incident numbers in `ages` that is not in `df` and 1558 incident numbers in `df` that is not in `ages`. Therefore, I will choose to ignore the age piece right now.
# only_ins <- df %>% select(incident_number) %>% unique()
# only_age_ins <- ages %>% select(incident_number) %>% unique()
# anti_join(only_age_ins, only_ins)

# Read in additional ages
ages <- read_xlsx("Additional IDOT info.xlsx") %>%
  clean_names() %>%
  rename("incident_number" = "incident") %>%
  select(incident_number, age_at_contact) %>%
  unique()

# Join for ages
tstops <- left_join(df, ages, by = "incident_number") %>%
  rename("age_at_contact" = "age_at_contact.y") %>%
  mutate(age_at_contact = case_when(is.na(age_at_contact.x) ~ age_at_contact,
                                    TRUE ~ age_at_contact.x)) %>%
  select(-age_at_contact.x)


# Below, I explored the differences in the two datasets, since joining seems to yield some extra data points. It seems there are 130 incident numbers in `ages` that is not in `df` and 1558 incident numbers in `df` that is not in `ages`. Therefore, I will choose to ignore the age piece right now.

 only_ins <- df %>% select(incident_number) %>% unique()
 only_age_ins <- ages %>% select(incident_number) %>% unique()
 anti_join(only_age_ins, only_ins)

```

```{r cleaning_df, echo=FALSE}
tstops <- df %>%
  group_by(incident_number) %>%
  # Create new variables that collapse multiple offenses into one column
  mutate(crime_descs  = paste(crime_description, collapse =", "),
         crime_codes = paste(crime_code, collapse = ", "),
         crime_status_descs = paste(crime_status_description, collapse = ", "),
         crime_categories = paste(crime_category, collapse = ", "),
         crime_category_descs = paste(crime_category_description, collapse = ", "),
         # Create new variable to count the number of offenses
         number_of_crimes = length(crime_codes)) %>%
  # ungroup
  ungroup() %>%
  # Deselect duplicate columns
  select(-crime_code, -crime_description, -crime_status_description, -crime_category, -crime_category_description) %>%
  # Filter for unique incidents
  unique()

## Create a new dataset that contains demographic information

tstops_demo <- tstops %>%
  # Select only for useful demographic and logistical information
  select(date_of_contact, time_of_contact, incident_number,
        motivation_for_stop_desc, race_description, sex_description,
        age_at_contact, residency_code, reason_for_stop,
        type_of_violation, vehicle_year, crime_descs, crime_status_descs, crime_category_descs) %>%
  # Extract year and month from date
  mutate(year = year(date_of_contact),
         month = month(date_of_contact),
         # Change "Caucasian" to "White" and factorize race
         race_description = as.factor(case_when(
           race_description == "CAUCASIAN" ~ "WHITE",
           race_description == "AFRICAN AMERICAN" ~ "BLACK",
           TRUE ~ race_description))) %>%
  filter(year != 2020)
```

## Disparity Ratios
According to the Indiana Department of Transportation 2018 annual report, disparity ratios are useful metrics for calculating the number of traffic stops that exceed the number of expected stops predicted from the population distribution.^[Insert citation here] Below, find the disparity ratio calculations by racial group for 2012-2020. 

```{r disparity-ratios, fig.height=3, fig.width=8, fig.cap="Disparity ratios by race by year"}
disparity_ratios <- tstops_demo %>%
  filter(!race_description %in% c("AMERICAN INDIAN/ALASKAN", "UNKNOWN")) %>%
  group_by(year) %>%
  mutate(all = n()) %>%
  ungroup() %>%
  group_by(year, race_description) %>%
  mutate(race_ct = n(),
         race_percentage = 100 * race_ct / all,
         disparity_ratio = case_when(
           year >= 2018 & race_description == "ASIAN" ~ race_percentage / 8.37,
           year >= 2018 & race_description == "BLACK" ~ race_percentage / 22.26,
           year >= 2018 & race_description == "HISPANIC" ~ race_percentage / 4.84,
           year >= 2018 & race_description == "WHITE" ~ race_percentage / 62.23,
           year == 2017 & race_description == "ASIAN" ~ race_percentage / 9.07,
           year == 2017 & race_description == "BLACK" ~ race_percentage / 20.36,
           year == 2017 & race_description == "HISPANIC" ~ race_percentage / 4.4,
           year == 2017 & race_description == "WHITE" ~ race_percentage / 63.7,
           year == 2016 & race_description == "ASIAN" ~ race_percentage / 9.91,
           year == 2016 & race_description == "BLACK" ~ race_percentage / 18.56,
           year == 2016 & race_description == "HISPANIC" ~ race_percentage / 4.04,
           year == 2016 & race_description == "WHITE" ~ race_percentage / 65.54,
           year <= 2015 & race_description == "ASIAN" ~ race_percentage / 9.96,
           year <= 2015 & race_description == "BLACK" ~ race_percentage / 18.07,
           year <= 2015 & race_description == "HISPANIC" ~ race_percentage / 3.52,
           year <= 2015 & race_description == "WHITE" ~ race_percentage / 67.26)) %>%
  ungroup() %>%
  group_by(year, race_description, sex_description) %>%
  mutate(race_gender_ct = n(),
         race_gender_percentage = 100 * race_gender_ct / all,
         disparity_ratio_gender = case_when(
           year >= 2018 & race_description == "ASIAN" ~ race_gender_percentage / 8.37,
           year >= 2018 & race_description == "BLACK" ~ race_gender_percentage / 22.26,
           year >= 2018 & race_description == "HISPANIC" ~ race_gender_percentage / 4.84,
           year >= 2018 & race_description == "WHITE" ~ race_gender_percentage / 62.23,
           year == 2017 & race_description == "ASIAN" ~ race_gender_percentage / 9.07,
           year == 2017 & race_description == "BLACK" ~ race_gender_percentage / 20.36,
           year == 2017 & race_description == "HISPANIC" ~ race_gender_percentage / 4.4,
           year == 2017 & race_description == "WHITE" ~ race_gender_percentage / 63.7,
           year == 2016 & race_description == "ASIAN" ~ race_gender_percentage / 9.91,
           year == 2016 & race_description == "BLACK" ~ race_gender_percentage / 18.56,
           year == 2016 & race_description == "HISPANIC" ~ race_gender_percentage / 4.04,
           year == 2016 & race_description == "WHITE" ~ race_gender_percentage / 65.54,
           year <= 2015 & race_description == "ASIAN" ~ race_gender_percentage / 9.96,
           year <= 2015 & race_description == "BLACK" ~ race_gender_percentage / 18.07,
           year <= 2015 & race_description == "HISPANIC" ~ race_gender_percentage / 3.52,
           year <= 2015 & race_description == "WHITE" ~ race_gender_percentage / 67.26)) %>%
  ungroup() %>%
  select(year, race_description, disparity_ratio, disparity_ratio_gender, sex_description, race_percentage) %>%
  unique()

## Only 2012 is off compared to the 2018 report, so I omitted it

# Graph disparity ratios
disparity_ratios %>%
  ggplot(aes(year, disparity_ratio, color = race_description)) +
  geom_line() +
  scale_x_continuous(breaks = c(2013, 2014, 2015, 2016,2017, 2018, 2019, 2020)) +
  scale_y_continuous(breaks = seq(0, 2.5, by = 0.5)) +
  theme_minimal() +
  labs(title = "Disparity Ratios",
       caption = "Data Publicly Available from Urbana Police Department") +
  scale_color_brewer(palette = "Dark2") +
  xlab("Year") +
  ylab("Disparity Ratio") +
  geom_hline(yintercept = 1, color = "grey")
```

Accounting for the interaction between gender and race, we get:

```{r message=FALSE, fig.height=4.5, fig.width=8, fig.cap="Disparity ratios by gender and race by year"}
# Graph disparity ratios by gender
disparity_ratios %>%
  ggplot(aes(year, disparity_ratio_gender, color = sex_description)) +
  geom_line() +
  scale_x_continuous(breaks = c(2013, 2014, 2015, 2016,2017, 2018, 2019, 2020)) +
  scale_y_continuous(breaks = seq(0, 2.5, by = 0.5)) +
  theme_minimal() +
  labs(title = "Disparity Ratios by Gender",
       caption = "Data Publicly Available from Urbana Police Department") +
  geom_hline(yintercept = 0.5, color = "grey") +
  facet_wrap(~race_description)

## To do gender*race disparity ratios: We'd need to either 1. Account for the different frequencies in male vs females being stopped, or 2. Find out the average # males and females stopped in Urbana to use as a "checkpoint" or baseline. The above chart displays disparity ratios by gender, which assumes that there are an equal number of men and women driving on the streets.
```

## Age Distributions

```{r age, fig.height=8, fig.width=10, message=FALSE, warning=FALSE, fig.cap="Age distributions of stopped drivers by race"}
# Age distribution of drivers in 2018, 2019, and 2020

tstops_demo %>%
  filter(!race_description %in% c("AMERICAN INDIAN/ALASKAN", "UNKNOWN"),
         !is.na(age_at_contact)) %>%
  ggplot(aes(x = age_at_contact, y = race_description)) +
  geom_density_ridges(show.legend = FALSE) +
  facet_wrap(~year) +
  labs(title = "Age Distribution of Drivers by Race",
       caption = "Data Publicly Available from the Urbana Police Department",
       xlab = "Age at Contact") +
  ylab(NULL)
```

```{r echo=FALSE, fig.height=4, fig.width=5, fig.cap="Age distribution of drivers in accidents"}
accidents <- accidents %>%
  mutate(year = year(Date)) %>%
  filter(!is.na(Age))

# Plot of age distribution
vertical <- accidents %>%
  filter(Race %in% c("W", "H", "A", "B")) %>%
  summarize(vertical = mean(Age))

vertical <- as.numeric(vertical)

accidents %>%
  filter(Race %in% c("W", "H", "A", "B")) %>%
  ggplot(aes(x = Age, y = Race)) +
  geom_density_ridges(show.legend = FALSE) +
  scale_y_discrete(labels = rev(c("White", "Hispanic", "Black", "Asian"))) +
  scale_x_continuous(breaks = seq(0, 100, by = 10)) +
  geom_vline(xintercept = vertical) +
  labs(title = "Age Distribution of Drivers Involved in Accidents",
       subtitle = "Mean age is about 39 years old")

# Plotting age distribution by at fault
accidents %>%
  filter(Race %in% c("W", "H", "A", "B")) %>%
  ggplot(aes(x = Age, y = Race)) +
  geom_density_ridges(show.legend = FALSE) +
  facet_wrap(~`At Fault`) +
  scale_y_discrete(labels = rev(c("White", "Hispanic", "Black", "Asian"))) +
  scale_x_continuous(breaks = seq(0, 100, by = 10)) +
  labs(title = "Age Distribution of Drivers Involved in Accidents",
       subtitle = "Faceted by whether the driver was at fault or not")
```

## Duration of Stops by Race

```{r duration, echo=FALSE, fig.height=4, fig.width=5}

# Join duration data with tstop demographic information
duration_join <- left_join(tstops_demo, durations, by = "incident_number") %>%
  filter(!race_description %in% c("AMERICAN INDIAN/ALASKAN", "UNKNOWN"),
         !is.na(duration_of_stop)) %>%
  group_by(year, race_description) %>%
  mutate(avg_duration = mean(duration_of_stop)) %>%
  ungroup() %>%
  mutate(race_description = fct_relevel(race_description, c("WHITE", "ASIAN", "BLACK", "HISPANIC")))

# Create barplot
duration_by_year <- duration_join %>%
  ggplot(aes(x = year, y = avg_duration ,fill = race_description)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(fill = "Race",
       title = "Average Duration of Stops by Race",
       subtitle = "Urbana, IL 2012-2017",
       caption = "Data Publicly Available from Urbana Police Department") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  scale_fill_brewer(palette = "PRGn") +
  xlab(NULL) +
  ylab("Duration in minutes")

duration_by_year
```

```{r stops_and_accidents, echo=FALSE, message=FALSE}

# After running quantile(accidents$Age), I've determined the following age
# groups: 0-24, 25-35, 36-52, and 53-99 that each represent a quantile and am
# transforming the accidents dataset as appropriate

accidents_yearly <- accidents %>%
  filter(Race %in% c("W", "H", "A", "B")) %>%
  mutate(Age = case_when(Age <= 24 ~ 1,
                         Age > 24 & Age <= 35 ~ 2,
                         Age > 35 & Age <= 52 ~ 3,
                         Age > 51 ~ 4),
         Race = case_when(Race == "W" ~ "WHITE",
                          Race == "H" ~ "HISPANIC",
                          Race == "A" ~ "ASIAN",
                          Race == "B" ~ "BLACK")) %>%
  group_by(year, Age, Race) %>%
  mutate(accidents_per_year = n()) %>% 
  ungroup() %>%
  select(year, accidents_per_year, Race, Age) %>%
  unique()

# Also make sure the factor levels in tstops_demo is the same as accidents so
# they can be joined

stops_yearly <- tstops_demo %>%
  filter(!race_description %in% c("AMERICAN INDIAN/ALASKAN", "UNKNOWN"),
         !is.na(age_at_contact),
         year >= 2014 & year <= 2019) %>%
  mutate(Age = case_when(age_at_contact <= 24 ~ 1,
                         age_at_contact > 24 & age_at_contact <= 35 ~ 2,
                         age_at_contact > 35 & age_at_contact <= 52 ~ 3,
                         age_at_contact > 51 ~ 4),
         Race = race_description) %>%
  group_by(year, Age, Race) %>%
  mutate(stops_per_year = n()) %>%
  ungroup() %>%
  select(year, stops_per_year, Race, Age) %>%
  unique()

# Join accidents and traffic stops data
stopsxaccidents <- left_join(stops_yearly, accidents_yearly, by = c("year", "Race", "Age"))

# General plot
ggplot(data = stopsxaccidents, mapping = aes(y = stops_per_year, x = accidents_per_year, color=Race)) +
  geom_point() +
  labs(title = "Stops per year")

stopsxaccidents %>%
  mutate(Race = fct_relevel(Race, c("WHITE", "ASIAN", "HISPANIC", "BLACK"))) %>%
  mutate(sa_ratio = stops_per_year / accidents_per_year) %>%
  ggplot(aes(x= year, y = sa_ratio, fill = Race)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~Age) +
  theme_minimal() +
  labs(title = "Ratio of Stops and Accidents",
       subtitle = "By age group and year")
```

```{r echo=FALSE, fig.height=6, fig.width=8}
# Create dataframe of outcome of traffic stops
outcomes <- tstops_demo %>% 
  select(race_description, sex_description, age_at_contact, incident_number, year) %>%
  mutate(outcome = case_when(str_detect(string = incident_number, pattern = "UT") ~ "Traffic Ticket",
                             str_detect(string = incident_number, pattern = "UW") ~ "Warning",
                             str_detect(string = incident_number, pattern = "UU") ~ "More Serious than Traffic Warning")) %>%
  filter(!race_description %in% c("AMERICAN INDIAN/ALASKAN", "UNKNOWN"))

# Analyze outcomes by race
outcomes %>%
  group_by(year, race_description) %>%
  mutate(all = n()) %>%
  ungroup() %>%
  group_by(year, race_description, outcome) %>%
  mutate(num_outcome = n(),
         prop_outcome = num_outcome / all) %>%
  ungroup() %>%
  select(race_description, year, outcome, prop_outcome) %>%
  ggplot(aes(x = year, y = prop_outcome, color=outcome)) +
  geom_line() +
  facet_wrap(~race_description) +
  theme(legend.position = "bottom") +
  xlab(NULL) +
  ylab("Proportion of outcome out of all outcomes") +
  labs(title = "Traffic Stop Outcomes by Race",
       subtitle = "Proportions of tickets, warnings, and more serious outcomes by race")
```

```{r echo=FALSE, fig.height = 8, fig.width = 8}
# Analyze outcomes by race
outcomes %>%
  group_by(year, race_description) %>%
  mutate(all = n()) %>%
  ungroup() %>%
  group_by(year, race_description, sex_description, outcome) %>%
  mutate(num_outcome = n(),
         prop_outcome = num_outcome / all) %>%
  ungroup() %>%
  select(race_description, sex_description, year, outcome, prop_outcome) %>%
  ggplot(aes(x = year, y = prop_outcome, color=outcome, linetype = sex_description)) +
  geom_line() +
  facet_grid(rows=vars(race_description)) +
  theme(legend.position = "bottom") +
  xlab(NULL) +
  ylab("Proportion of outcome out of all outcomes") +
  labs(title = "Traffic Stop Outcomes by Race",
       subtitle = "Proportions of tickets, warnings, and more serious outcomes by race and gender")
```


Discuss: vertical vs horizontal variation for scatterplots and what that means; younger folks and ratios of being stopped vs being in accidents compared to the general trends of the age of drivers involved in stops; disparity ratios across gender/race and other factors (time of day?)

## Methodology

## Appendix 

## References

## Take out 2020, take out ugly line graph of age
- UT: traffic ticket/citation
- UW: warning - warning ticket
- UU: more serious than a warning (DUI, warrant, etc) - a report had to be written for it.
